Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-04-05 Ben_Manuscripts/thesis/structure.tex
Updates to outline and minor changes to figures

  More highly selective nanoporous membranes would be extremely useful for
  performing complex aqueous separations with seawater and various types of
  wastewater. For example, Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
  represent just a few of the diverse contaminants of water sources. By
  efficiently separating contaminants from feed solutions with highly selective
  membranes, it is possible to reduce the number of required membrane passes and
  post-treatment steps needed for a given filtration process
  \cite{werber_materials_2016}, thus lowering energy requirements. Additionally,
  one can also extract valuable resources from the feed streams. For example,
  flowback water produced during hydraulic fracturing of shale formations
  contains dissolved species such as acetate whose extraction has economic value
  \cite{dischinger_application_2017}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, including ions. However, commercially available amorphous
  RO membranes and porous NF membranes lack uniformity, which severely impacts
  their selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures that offer separation pathways that are tortuous and polydisperse in
  size \cite{werber_materials_2016}. Tortuosity increases the effective length
  that a solute must travel, while pore size polydispersity limits membrane
  selectivity.

  Ordered, nanostructured porous membrane materials have become an increasingly popular
  research area for aqueous separation applications because they offer the
  ability to control pore architecture at the molecular scale, thereby
  permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. However, their development has been limited
  by the ability to synthesize and scale various technologies. Graphene sheets
  are atomically thick, which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}.  Molecular dynamics (MD) simulations of
  carbon nanotubes have shown promise \cite{humplik_nanostructured_2011}, but
  synthetic techniques are unable to achieve scalable alignment and pore
  monodispersity \cite{hata_water-assisted_2004,maruyama_growth_2005}.
  Consequently, there is a need for new materials that could provide scalable
  nanostructured porous membranes.

  Preliminary evidence has shown that cross-linked lyotropic liquid crystal
  (LLC) membranes can be produced at moderate scale and may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} and can be cross-linked to create mechanically strong
  membrane films with periodic pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike the pores in most commercial NF membranes,
  polymeric LLC membrane pores are uniform in size because they are formed by
  self-assembly. Since LLC polymer membranes lack an appreciable pore size
  distribution, they inherently exhibit high selectivity due to their strict
  molecular weight cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the
  LLC monomers examined in this paper are salts, and therefore lead to Donnan
  exclusion of ions in solution. The membrane gains a net surface charge when
  counterions from the head groups that line the pore walls escape into the feed
  solution in an effort to balance the gradients of concentration and electric
  potential \cite{donnan_theory_1995}.

  The feasibility of nanostructured LLC polymer membranes for selective
  separations has been demonstrated using LLC monomers that form the type 1
  bicontinuous cubic (Q\textsubscript{I})
  \cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012}
  and the inverted hexagonal (H\textsubscript{II}) \cite{zhou_supported_2005}
  phases. When separating organic solutes from NaCl, Q\textsubscript{I}-phase
  membrane filtration experiments have shown selectivity 2--3 times higher than
  commercial RO and 6--12 times higher than commercial NF membranes
  \cite{dischinger_application_2017}. When separating a series of various sized
  dyes, the H\textsubscript{II}-phase membrane showed complete rejection of dyes
  bigger than 1.2 nm in size \cite{zhou_supported_2005}.

  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:hex_packing_simple}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  ca.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in the literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. Q\textsubscript{I}-phase membranes consist of
  a tortuous network of three dimensionally interconnected pores that prevent
  optimal through-plane transport. In contrast, the densely packed, non-tortuous
  and uniform sized pores of H\textsubscript{II}-phase membranes represent the
  ideal geometry for achieving high solute
  flux\cite{matyka_tortuosity-porosity_2008}. However, the hexagonally packed LC
  domains of the H\textsubscript{II}-phase generally form mutually unaligned
  domains, which hurts membrane permeability. This domain scale misalignment had
  inhibited further development of this technology, and research efforts were
  focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{zhou_new_2007}.

  However, recently researchers have learned how to macroscopically align the
  hexagonal domains which has revived research into H\textsubscript{II}-phase LLC
  polymer membranes. In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
  domains, created by the ``dry" monomer Na-GA3C11, using a magnetic field
  with subsequent cross-linking to lock the structure in
  place\cite{feng_scalable_2014}. In 2016, Feng et al.~showed that one could
  also obtain the same result by confining the neat monomer between PDMS or glass
  substrates since hexagonal mesophases preferentially anchor perpendicular to
  both surfaces\cite{feng_thin_2016}. Current experimental efforts are focused
  on extending the method to the H\textsubscript{II} phase and characterizing
  the performance of these newly aligned systems.

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{columns.pdf}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order, then assemble into pores with 
	hydrophilic head groups (blue) facing towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

@@

  % BJC: reworked text
  \label{rework:purpose}
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our understanding of the role
  of monomer structure in solute transport and provide guidelines to reduce the 
  large chemical space available to design monomers for creation of 
  separation-specific membranes. It is important that we model the system with
  a nanoscopic pore structure that is maximally consistent with experimental data
  in order to best approximate the chemical environment experienced by solutes
  in future simulated transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  In this paper, we perform molecular modeling of the Col\textsubscript{h}
  thermotropic (i.e., solvent-free) assembly formed by Na-GA3C11. Compared to the
  lyotropic (i.e, solvent-containing) H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point. The system is not
  assembled in aqueous phase which allowed us to simulate longer timescales by
  omitting the solvent in the simulations, and there exists detailed experimental
  characterization of the fully aligned state, including 2D-WAXS patterns
  (Figure~\ref{fig:WAXS}) that are useful for reconstructing structural data.

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction. Section \ref{S-section:parameterization} 
  of the SI provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  that a united-atom model of an LLC can self-assemble into the
  H\textsubscript{II} phase configuration in $\sim$1000 ns
  \cite{mondal_self-assembly_2013}. We attempted atomistic self-assembly by
  packing monomers into a box using Packmol \cite{martinez_packmol:_2009}.
  Simulations of greater than 100 ns show no indicators of progress towards an
  ordered system (see Section \ref{S-section:self_assembly} of the SI). To bypass
  the slow self-assembly process, we used automated procedures to
  assemble monomers into a structure close to one of a number of hypothesized
  equilibrium configurations (See Figure \ref{S-fig:build_procedure} of the SI).
  These initial structures may result in equilibration into slowly-intercoverting
  metastable states, an issue we will address later in the paper.

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number of stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to minimize finite size effects with reasonable computational expense 
  and to obtain sufficient resolution when simulating XRD patterns (see further discussion
  in Section \ref{S-section:monomers_per_column} of the SI).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the
  equilibrium configuration is sensitive to some while insensitive to others. For
  example, we find the starting pore radius, defined as the distance of a chosen
  head group carbon from the pore's central axis, does not influence the
  equilibrium structure if one chooses a reasonable value. The initial distance
  between pores, within a wide range, also has little effect on the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose a pore radius of 0.5 nm
  and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm for our initial configurations. A sensitivity analysis of both
  parameters is presented in the SI, Section
  \ref{S-section:initial_config_dependence}. The distance between vertically
  stacked monomers, the $xy$ position of monomers with respect to vertically
  adjacent monomers, and the number of columns-per-pore do influence the
  equilibrated structure and require further justification for their choices. We
  rely on experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration
  based on R-$\pi$ and then allowed the system to readjust during equilibration.
  We rotated each monomer so the plane of its aromatic head group would be
  coplanar with the $xy$ plane. We explored three different initial monomer
  spacings. The first is exactly equal to R-$\pi$ with monomers placed so
  aromatic rings stack 3.7 \AA~apart in the $z$-direction. We also extensively
  explored a second system with an initial spacing of 5 \AA. We briefly explored
  a third system with an initial spacing of 10 \AA. However this largest spacing
  yields non-physical behavior which is detailed in the SI, Section
  \ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns-per-pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

@@

  We developed equilibration schemes for creating dry (i.e. thermotropic LC phase)
  and wet (i.e. LLC phase) configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat, followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in Section \ref{S-section:equilibration} of the SI.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of
  sodium ions in their respective pores. We generated pore spacing statistics
  using the bootstrapping technique (See Section \ref{S-section:p2p_stats} of the
  SI). The pore spacings calculated in this way are consistent with one half of
  the $x$ and $y$ box vectors (See Table \ref{S-table:p2p} in the SI). However,
  our method for their calculation does a better job capturing the spread of
  pore-to-pore distances.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D-WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D-SAXS which was generated from the same
  sample (Figure \ref{S-fig:2DSAXS} of the SI). Since the d\textsubscript{200} peak is
  partially exposed in the experimental 2D-WAXS pattern, we normalized the
  experimental 2D-SAXS cross-section by matching the intensity of the
  d\textsubscript{200} peak between it and the experimental 2D-WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D-WAXS
  pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  We explored the pores' compositions by measuring the average number densities
  of various monomer components as a function of distance from the pore centers.
  We looked at the average number density of sodium ions, aromatic rings and 
  carbon atoms making up the monomer tails. We binned the radial distance of all
  atoms in each group from the pore centers, then normalized by the volume of the
  annulus defined by the bin edges and the z box vector (See Figure \ref{S-fig:rdf_diagram}
  of the SI).

@@

  We calculated ionic conductivity using the Nernst-Einstein relationship, which 
  relates the DC ionic conductivity, $\sigma$, to ion diffusivity, $D$, 
  concentration, $C$, ion charge, $q$, the Boltzmann constant, $k_b$, and 
  absolute temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the $z$-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell. 
  More details are provided in the SI, Section \ref{S-section:ionic_conductivity}.

@@

  We based our cross-linking algorithm on the known reaction mechanism 
  (Figure \ref{S-fig:xlink_mech} of the SI). The reaction 
  takes place at the terminal vinyl groups
  on each alkane tail. The procedure is carried out iteratively. Each iteration, the
  algorithm chooses carbons to cross-link based on the distance between eligible 
  carbon pairs. The algorithm then updates the topology with the new bonds and atom
  types, energy minimizes the system and runs a short simulation before selecting 
  the next group of eligible carbons atoms.

>>>>

%  More highly selective nanoporous membranes would be extremely useful for
%  performing complex aqueous separations with seawater and various types of
%  wastewater. For example, Sodium chloride and boron in seawater
%  \cite{fritzmann_state---art_2007} and organic micropollutants found in
%  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
%  represent just a few of the diverse contaminants of water sources. By
%  efficiently separating contaminants from feed solutions with highly selective
%  membranes, it is possible to reduce the number of required membrane passes and
%  post-treatment steps needed for a given filtration process
%  \cite{werber_materials_2016}, thus lowering energy requirements. Additionally,
%  one can also extract valuable resources from the feed streams. For example,
%  flowback water produced during hydraulic fracturing of shale formations
%  contains dissolved species such as acetate whose extraction has economic value
%  \cite{dischinger_application_2017}.  
%
%  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
%  filtration processes that can be used to separate solutes on the order of 1 nm
%  in size and smaller, including ions. However, commercially available amorphous
%  RO membranes and porous NF membranes lack uniformity, which severely impacts
%  their selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
%  fabrication involves the spontaneous assembly of polymers into disordered
%  structures that offer separation pathways that are tortuous and polydisperse in
%  size \cite{werber_materials_2016}. Tortuosity increases the effective length
%  that a solute must travel, while pore size polydispersity limits membrane
%  selectivity.
%
%  Ordered, nanostructured porous membrane materials have become an increasingly popular
%  research area for aqueous separation applications because they offer the
%  ability to control pore architecture at the molecular scale, thereby
%  permitting the design of solute-specific separation membranes
%  \cite{humplik_nanostructured_2011}. However, their development has been limited
%  by the ability to synthesize and scale various technologies. Graphene sheets
%  are atomically thick, which results in excellent water permeability but defects
%  during manufacturing severely impact selectivity
%  \cite{cohen-tanugi_multilayer_2016}.  Molecular dynamics (MD) simulations of
%  carbon nanotubes have shown promise \cite{humplik_nanostructured_2011}, but
%  synthetic techniques are unable to achieve scalable alignment and pore
%  monodispersity \cite{hata_water-assisted_2004,maruyama_growth_2005}.
%  Consequently, there is a need for new materials that could provide scalable
%  nanostructured porous membranes. 
%
%  Preliminary evidence has shown that cross-linked lyotropic liquid crystal
%  (LLC) membranes can be produced at moderate scale and may be capable of
%  performing highly selective separations. LLCs are amphiphilic molecules that
%  have the ability to self-assemble into porous nanostructures
%  \cite{smith_ordered_1997} and can be cross-linked to create mechanically strong
%  membrane films with periodic pores on the order of 1 nm in diameter
%  \cite{zhou_supported_2005}. Unlike the pores in most commercial NF membranes,
%  polymeric LLC membrane pores are uniform in size because they are formed by
%  self-assembly. Since LLC polymer membranes lack an appreciable pore size
%  distribution, they inherently exhibit high selectivity due to their strict
%  molecular weight cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the
%  LLC monomers examined in this paper are salts, and therefore lead to Donnan
%  exclusion of ions in solution. The membrane gains a net surface charge when
%  counterions from the head groups that line the pore walls escape into the feed
%  solution in an effort to balance the gradients of concentration and electric
%  potential \cite{donnan_theory_1995}.    
%
%  The feasibility of nanostructured LLC polymer membranes for selective
%  separations has been demonstrated using LLC monomers that form the type 1
%  bicontinuous cubic (Q\textsubscript{I})
%  \cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012}
%  and the inverted hexagonal (H\textsubscript{II}) \cite{zhou_supported_2005}
%  phases. When separating organic solutes from NaCl, Q\textsubscript{I}-phase
%  membrane filtration experiments have shown selectivity 2--3 times higher than
%  commercial RO and 6--12 times higher than commercial NF membranes
%  \cite{dischinger_application_2017}. When separating a series of various sized
%  dyes, the H\textsubscript{II}-phase membrane showed complete rejection of dyes
%  bigger than 1.2 nm in size \cite{zhou_supported_2005}. 
%
%  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:hex_packing_simple}) has a
%  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
%  The H\textsubscript{II} phase forms at room temperature in the presence of
%  ca.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
%  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
%  form the same hexagonal columnar structure which, in the literature, has been
%  referred to as the Col\textsubscript{h} thermotropic
%  phase\cite{feng_scalable_2014}. Q\textsubscript{I}-phase membranes consist of
%  a tortuous network of three dimensionally interconnected pores that prevent
%  optimal through-plane transport. In contrast, the densely packed, non-tortuous
%  and uniform sized pores of H\textsubscript{II}-phase membranes represent the
%  ideal geometry for achieving high solute
%  flux\cite{matyka_tortuosity-porosity_2008}. However, the hexagonally packed LC
%  domains of the H\textsubscript{II}-phase generally form mutually unaligned
%  domains, which hurts membrane permeability. This domain scale misalignment had
%  inhibited further development of this technology, and research efforts were
%  focused on the Q\textsubscript{I} phase, whose geometry does not require
%  alignment~\cite{zhou_new_2007}. 
%
%  However, recently researchers have learned how to macroscopically align the
%  hexagonal domains which has revived research into H\textsubscript{II}-phase LLC
%  polymer membranes. In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
%  domains, created by the ``dry" monomer Na-GA3C11, using a magnetic field
%  with subsequent cross-linking to lock the structure in
%  place\cite{feng_scalable_2014}. In 2016, Feng et al.~showed that one could
%  also obtain the same result by confining the neat monomer between PDMS or glass
%  substrates since hexagonal mesophases preferentially anchor perpendicular to
%  both surfaces\cite{feng_thin_2016}. Current experimental efforts are focused
%  on extending the method to the H\textsubscript{II} phase and characterizing
%  the performance of these newly aligned systems.
%
%  \begin{figure}
%	\centering
%	\begin{subfigure}{.3\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
%		\caption{}~\label{fig:monomer}
%	\end{subfigure}
%	\begin{subfigure}{.3\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
%		\caption{}~\label{fig:atomistic_monomer}
%	\end{subfigure}
%	\begin{subfigure}{0.3\linewidth}
%		\centering
%		\includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
%		\caption{}~\label{fig:wedge}
%	\end{subfigure}
%		\begin{subfigure}{0.4\linewidth}
%		\centering
%		\includegraphics[width=0.6\textwidth]{columns.pdf}
%		\caption{}~\label{fig:wedge_layer}
%	\end{subfigure}
%	\begin{subfigure}{0.4\linewidth}
%		\centering
%		\includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
%		\caption{}~\label{fig:hex_packing_simple}
%	\end{subfigure}
%	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
%	exhibits wedge-like character. (d) Monomers stack on top of each other to
%	create columns with short range order, then assemble into pores with 
%	hydrophilic head groups (blue) facing towards the pore center. (e) The 
%	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
%  \end{figure}

@@

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our understanding of the role
  of monomer structure in solute transport and provide guidelines to reduce the 
  large chemical space available to design monomers for creation of 
  separation-specific membranes. It is important that we model the system with
  a nanoscopic pore structure that is maximally consistent with experimental data
  in order to best approximate the chemical environment experienced by solutes
  in future simulated transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  In this chapter, we perform molecular modeling of the Col\textsubscript{h}
  thermotropic (i.e., solvent-free) assembly formed by Na-GA3C11. Compared to the
  lyotropic (i.e, solvent-containing) H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point. The system is not
  assembled in aqueous phase which allowed us to simulate longer timescales by
  omitting the solvent in the simulations, and there exists detailed experimental
  characterization of the fully aligned state, including 2D-WAXS patterns
  (Figure~\ref{fig:WAXS}) that are useful for reconstructing structural data.

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction. Section \ref{S-section:parameterization} 
  of Appendix A provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  that a united-atom model of an LLC can self-assemble into the
  H\textsubscript{II} phase configuration in $\sim$1000 ns
  \cite{mondal_self-assembly_2013}. We attempted atomistic self-assembly by
  packing monomers into a box using Packmol \cite{martinez_packmol:_2009}.
  Simulations of greater than 100 ns show no indicators of progress towards an
  ordered system (see Section \ref{S-section:self_assembly} of Appendix A). To bypass
  the slow self-assembly process, we used automated procedures to
  assemble monomers into a structure close to one of a number of hypothesized
  equilibrium configurations (See Figure \ref{S-fig:build_procedure} of Appendix A).
  These initial structures may result in equilibration into slowly-intercoverting
  metastable states, an issue we will address later in the paper.

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number of stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to minimize finite size effects with reasonable computational expense 
  and to obtain sufficient resolution when simulating XRD patterns (see further discussion
  in Section \ref{S-section:monomers_per_column} of Appendix A).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the
  equilibrium configuration is sensitive to some while insensitive to others. For
  example, we find the starting pore radius, defined as the distance of a chosen
  head group carbon from the pore's central axis, does not influence the
  equilibrium structure if one chooses a reasonable value. The initial distance
  between pores, within a wide range, also has little effect on the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose a pore radius of 0.5 nm
  and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm for our initial configurations. A sensitivity analysis of both
  parameters is presented in Appendix A, Section
  \ref{S-section:initial_config_dependence}. The distance between vertically
  stacked monomers, the $xy$ position of monomers with respect to vertically
  adjacent monomers, and the number of columns-per-pore do influence the
  equilibrated structure and require further justification for their choices. We
  rely on experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration
  based on R-$\pi$ and then allowed the system to readjust during equilibration.
  We rotated each monomer so the plane of its aromatic head group would be
  coplanar with the $xy$ plane. We explored three different initial monomer
  spacings. The first is exactly equal to R-$\pi$ with monomers placed so
  aromatic rings stack 3.7 \AA~apart in the $z$-direction. We also extensively
  explored a second system with an initial spacing of 5 \AA. We briefly explored
  a third system with an initial spacing of 10 \AA. However this largest spacing
  yields non-physical behavior which is detailed in Appendix A, Section
  \ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns-per-pore. See Figure 
  \ref{S-fig:stacking} of Appendix A for a detailed illustration 
  of the initial configurations in each mode.

@@

  We developed equilibration schemes for creating dry (i.e. thermotropic LC phase)
  and wet (i.e. LLC phase) configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat, followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in Section \ref{S-section:equilibration} of Appendix A.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of
  sodium ions in their respective pores. We generated pore spacing statistics
  using the bootstrapping technique (See Section \ref{S-section:p2p_stats} of the
  SI). The pore spacings calculated in this way are consistent with one half of
  the $x$ and $y$ box vectors (See Table \ref{S-table:p2p} in Appendix A). However,
  our method for their calculation does a better job capturing the spread of
  pore-to-pore distances.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D-WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D-SAXS which was generated from the same
  sample (Figure \ref{S-fig:2DSAXS} of Appendix A). Since the d\textsubscript{200} peak is
  partially exposed in the experimental 2D-WAXS pattern, we normalized the
  experimental 2D-SAXS cross-section by matching the intensity of the
  d\textsubscript{200} peak between it and the experimental 2D-WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D-WAXS
  pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  We explored the pores' compositions by measuring the average number densities
  of various monomer components as a function of distance from the pore centers.
  We looked at the average number density of sodium ions, aromatic rings and 
  carbon atoms making up the monomer tails. We binned the radial distance of all
  atoms in each group from the pore centers, then normalized by the volume of the
  annulus defined by the bin edges and the z box vector (See Figure \ref{S-fig:rdf_diagram}
  of Appendix A).

@@

  We calculated ionic conductivity using the Nernst-Einstein relationship, which 
  relates the DC ionic conductivity, $\sigma$, to ion diffusivity, $D$, 
  concentration, $C$, ion charge, $q$, the Boltzmann constant, $k_b$, and 
  absolute temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the $z$-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell. 
  More details are provided in Appendix A, Section \ref{S-section:ionic_conductivity}.

@@

  We based our cross-linking algorithm on the known reaction mechanism 
  (Figure \ref{S-fig:xlink_mech} of Appendix A). The reaction 
  takes place at the terminal vinyl groups
  on each alkane tail. The procedure is carried out iteratively. Each iteration, the
  algorithm chooses carbons to cross-link based on the distance between eligible 
  carbon pairs. The algorithm then updates the topology with the new bonds and atom
  types, energy minimizes the system and runs a short simulation before selecting 
  the next group of eligible carbons atoms.
